---
title: "obj1"
output: html_document
date: "2024-07-24"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

# How do the composite indicators compare against each other?

```{r}
rq_packages <- c("tidyverse", "ggpubr", "table1")
 
installed_packages <- rq_packages %in% rownames(installed.packages())
if (any(installed_packages == FALSE)) {
  install.packages(rq_packages[!installed_packages])
}
 
lapply(rq_packages, require, character.only = T)
rm(list= c("rq_packages", "installed_packages"))


eth <- read.csv("eth_nar_pa.csv")
ind <- read.csv("ind_nar_pa.csv") 
nga <- read.csv("nga_nar_pa.csv")
```

## Distributions of MAR and MPA

```{r}

# MAR
# Distribution of MAR values in Ethiopia
ggplot(eth, aes(x=mar))+
  geom_histogram(fill = "#486DB8")+
  theme_minimal()+
  labs(x = "MAR", y = "count",  title = "Distribution of MAR values in Ethiopia")+
  theme_classic()



# Distribution of MAR values in India
ggplot(ind, aes(x=mar))+
  geom_histogram(fill =  "#486DB8")+
  theme_minimal()+
  labs(x = "MAR", y = "count", title = "Distribution of MAR values in India")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme_classic()


# Distribution of MAR values in Nigeria
ggplot(nga, aes(x=mar))+
  geom_histogram(fill = "#486DB8")+
  theme_minimal() +
  labs(x = "MAR", y = "count", title = "Distribution  of MAR values in Nigeria")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme_classic()

```

## Interpretation

The distribution of mar in every country is **left** skewed.

The majority of each sample has a high MAR value.

Ethiopia and Nigeria have very high counts of high MAR. Nigeria's stepwise distribution suggests a higher number of individuals with increasing MAR values. The majority of each population have diets that are relatively adequate, with many approaching full adequacy.


# How do the NAR distributions contribute to the MAR distribution?
```{r}

# to see the shape of the distributions, i changed the x limits so that truncated values are not included, and the y limits so that more wider distributions are clearer

# vitamin a 
ggplot(eth, aes(x=vita_rae_nar))+
  geom_histogram() +
  theme_minimal()+
  xlim(-0.05,1)+ 
  ylab("Count")+
  xlab("Vitamin A NAR")+
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme_classic()

# folate 
ggplot(eth, aes(x=folate_nar))+
  geom_histogram() +
  theme_minimal()+
  xlim(-0.05,1)+ 
  ylim(0, 2000)+
  ylab("Count")+
  xlab("Folate NAR")+
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme_classic()


# vitamin b12
ggplot(eth, aes(x=vitb12_nar))+
  geom_histogram() +
  theme_minimal()+
  xlim(-0.05,1)+ 
  ylab("Count")+
  xlab("Vitamin B12 NAR")+
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme_classic()

# iron
ggplot(eth, aes(x=fe_nar))+
  geom_histogram() +
  theme_minimal()+
  xlim(-0.05,1)+ 
    ylim(0, 1000)+
  ylab("Count")+
  xlab("Iron NAR")+
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme_classic()

# zinc
ggplot(eth, aes(x=zn_nar))+
  geom_histogram() +
  theme_minimal()+
  xlim(-0.05,1)+
    ylim(0, 1000)+
  ylab("Count")+
  xlab("Zinc NAR")+
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme_classic()

# thiamine
ggplot(eth, aes(x=thia_nar))+
  geom_histogram() +
  theme_minimal()+
  xlim(-0.05,1)+ 
    ylim(0, 1000)+
  ylab("Count")+
  xlab("Thiamine NAR")+
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme_classic()

# Riboflavin 
ggplot(eth, aes(x=ribo_nar))+
  geom_histogram() +
  theme_minimal()+
  xlim(-0.05,1)+
    ylim(0, 2000)+
  ylab("Count")+
  xlab("Riboflavin NAR")+
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme_classic()

# Niacin 
ggplot(eth, aes(x=niac_nar))+
  geom_histogram() +
  theme_minimal()+
  xlim(-0.05,1)+ 
    ylim(0, 1000)+
  ylab("Count")+
  xlab("Niacin NAR")+
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme_classic()

# vit b6 
ggplot(eth, aes(x=vitb6_nar))+
  geom_histogram() +
  theme_minimal()+
  xlim(-0.05,1)+ 
    ylim(0, 1000)+
  ylab("Count")+
  xlab("Vitamin B6 NAR")+
  theme(plot.title = element_text(hjust = 0.5))+ 
  theme_classic()
```



```{r}
# MPA 
# Distribution of MPA values in Ethiopia
ggplot(eth, aes(x=mpa))+
  geom_histogram(fill = "#8E48B8")+
  theme_minimal()+
  labs(x = "MPA", y = "count", title = "Distribution of MPA values in Ethiopia")+
  theme(plot.title = element_text(hjust = 0.5))+
  theme_classic()
  
# Distribution of MPA values in India
ggplot(ind, aes(x=mpa))+
  geom_histogram()+
  theme_minimal()+
  labs(x = "MPA", y = "count", title = "Distribution of MPA values in India")+
  theme(plot.title = element_text(hjust = 0.5))


# Distribution of MPA values in Nigeria
ggplot(nga, aes(x=mpa))+
  geom_histogram()+
  theme_minimal()+
  labs(x = "MPA", y = "count", title = "Distribution of MPA values in Nigeria")+
  theme(plot.title = element_text(hjust = 0.5))
```

## Interpretation

The MPA distributions are far less conventional with peaks and lows. The distributions show a lot of variation in the mean probability of adequacy across the three populations and suggest a degree of variability in meeting nutrient requirements

Ethiopia - More peaks counts for higher MPAs (closer to 1), meaning that on avergae, large numbers of the population are likely to meet nutrient needs.
Peaks at 0 and group of peaks in middle (similar pa values?)

India & Nigeria - high counts of low MPA (close to 0), decreasing counts for higher MPA. Majority of the population is unlikely to meet their requirement needs. <!-- this is in contrast to MAR - where the majority of the Nigerian population have MAR values close to 1. -->


# How do the individual PA distributions contribute to the MPA distribution?
```{r}

# plotting probability againt proportion of the population

# vitamin a 
ggplot(eth, aes(x=vita_rae_pa))+
 geom_histogram(aes(y = ..count../sum(..count..)))+
  ylim(0,1)+
  theme_minimal()+
  labs(x = "Vitamin A PA", y = "Proportion", title = "Distribution of Vitamin A PA")+
  theme_classic()


# folate 
ggplot(eth, aes(x=folate_pa))+
 geom_histogram(aes(y = ..count../sum(..count..)))+
  ylim(0,1)+
  theme_minimal()+
  labs(x = "Folate PA", y = "Proportion")+
  theme_classic()


# vitamin b12 
ggplot(eth, aes(x=vitb12_pa))+
 geom_histogram(aes(y = ..count../sum(..count..)))+
  ylim(0,1)+
  theme_minimal()+
  labs(x = "Vitamin B12 PA", y = "Proportion")+
  theme_classic()

# iron 
ggplot(eth, aes(x=fe_pa))+
 geom_histogram(aes(y = ..count../sum(..count..)))+
  ylim(0,1)+
  theme_minimal()+
  labs(x = "Iron PA", y = "Proportion")+
  theme_classic()

# zinc 
ggplot(eth, aes(x=zn_pa))+
 geom_histogram(aes(y = ..count../sum(..count..)))+
  ylim(0,1)+
  theme_minimal()+
  labs(x = "Zinc PA", y = "Proportion")+
  theme_classic()

# thiamine 
ggplot(eth, aes(x=thia_pa))+
 geom_histogram(aes(y = ..count../sum(..count..)))+
  ylim(0,1)+
  theme_minimal()+
  labs(x = "Thiamine PA", y = "Proportion")+
  theme_classic()


# riboflavin 
ggplot(eth, aes(x=ribo_pa))+
 geom_histogram(aes(y = ..count../sum(..count..)))+
  ylim(0,1)+
  theme_minimal()+
  labs(x = "Riboflavin PA", y = "Proportion")+
  theme_classic()

# niacin 
ggplot(eth, aes(x=niac_pa))+
 geom_histogram(aes(y = ..count../sum(..count..)))+
  ylim(0,1)+
  theme_minimal()+
  labs(x = "Niacin PA", y = "Proportion", title = "Distribution of Niacin PA")+
  theme_classic()

# vitamin b6 
ggplot(eth, aes(x=vitb6_pa))+
 geom_histogram(aes(y = ..count../sum(..count..)))+
  ylim(0,1)+
  theme_minimal()+
  labs(x = "Vitamin B6 PA", y = "Proportion")+
  theme_classic()

```



```{r}
## Comparing Descriptive (Tables)


label(eth$mar) <- "Mean Adequacy Ratio"
label(eth$mpa) <- "Mean Probability of Adequacy"
# table1(~mar + mpa | res, data = eth, na.rm = TRUE, digits = 3, format.number = TRUE)

label(ind$mar) <- "Mean Adequacy Ratio"
label(ind$mpa) <- "Mean Probability of Adequacy"
# table1(~mar + mpa | res, data = ind, na.rm = TRUE, digits = 3, format.number = TRUE)

label(nga$mar) <- "Mean Adequacy Ratio"
label(nga$mpa) <- "Mean Probability of Adequacy"
# table1(~mar + mpa | res, data = nga, na.rm = TRUE, digits = 3, format.numebr = TRUE)

# Focusing on median because of skewed distributions 

# Overall
  # MAR higher than MPA
  # MAR does not vary between countries, however the MPA shows clear differences (e.g. Ethiopia standing out), as reflected in distributions^

# Stratifying sub nationally 
  # Both higher in urban than rural 
  # ethiopia?
```

```{r}
ethh <- eth %>% 
  mutate(country = "Ethiopia")
indd <- ind %>% 
  mutate(country = "India")
ngaa <- nga %>% 
  mutate(country ="Nigeria")

all <- rbind(ethh, indd, ngaa) %>% drop_na()

# so that rural is one group
all <- all %>%
  mutate(res = str_trim(res)) %>%  
  mutate(res = tolower(res)) %>%  
  mutate(res = recode(res, 'rural' = 'Rural', 'urban' = 'Urban')) 


summary <- function(x, ...) {
  x <- na.omit(x)
  
  mean_val <- mean(x, na.rm = FALSE) %>% round(3)
  sd_val <- sd(x, na.rm = FALSE) %>% round(3)
  q1 <- quantile(x, 0.25) %>% round(3)
  med <- median(x, na.rm = FALSE) %>%  round(3)
  q3 <- quantile(x, 0.75) %>% round(3)
  iqr <- q3 - q1
  
  # mean (SD)
  mean_sd <- paste0(mean_val, " (", sd_val, ")")
  
  # median [Q1 - Q3]
  median_iqr <- paste0(med, " [", iqr, "]")
  
  c("", 'Mean (SD)' = mean_sd, 'Median [IQR]' = median_iqr)
}

table1(~ mar + mpa | country, 
       data = all, 
       na.rm = FALSE, 
       render.continuous = summary,
       digits = 3, 
       format.number = TRUE)

table1(~ mar + mpa | country*res, 
       data = all, 
       na.rm = FALSE, 
       render.continuous = summary,
       digits = 3, 
       format.number = TRUE)

table1(~ mar + mpa | country*sep_quintile, 
       data = all, 
       na.rm = FALSE, 
       render.continuous = summary,
       digits = 3, 
       format.number = TRUE)


# steady increase in mar and mpa as sep_quintile increases (as population gets richer)

```

# Plotting MAR and MPA against each other

```{r}
# Ethiopia 
mar_mpa_e <- ggscatter(eth, x = "mar",
                       y = "mpa",
                        alpha = 0.1,
                       conf.int = TRUE )
mar_mpa_e + stat_cor(method = "spearman",
                p.accuracy = 0.001,
                r.accuracy = 0.01)+
  labs(x = "Mean Adequacy Ratio", y = "Mean Probability of Adequacy", title = "Ethiopia")+
  xlim(0, 1)+
  ylim(0, 1)


#India 
mar_mpa_i <- ggscatter(ind, x = "mar", y = "mpa",
                       alpha = 0.1,
                       conf.int = TRUE )
mar_mpa_i + stat_cor(method = "spearman",
                p.accuracy = 0.001,
                r.accuracy = 0.01)+
  labs(x = "Mean Adequacy Ratio", y = "Mean Probability of Adequacy", title = "India")+
  xlim(0, 1)+
  ylim(0, 1)

#Nigeria
mar_mpa_n <- ggscatter(nga, x = "mar",
                       y = "mpa",
                       alpha = 0.1,
                       conf.int = TRUE )
mar_mpa_n + stat_cor(method = "spearman",
                p.accuracy = 0.001,
                r.accuracy = 0.01)+
  labs(x = "Mean Adequacy Ratio", y = "Mean Probability of Adequacy", title = "Nigeria")+
  xlim(0, 1)+
  ylim(0, 1)

# All have very high correlations, suggesting that MAR and MPA are closely aligned - the higher the mean adequacy ratio, the higher the likelihood of meeting nutritional requirements 
```

# How do they compare against each other on a very controlled sample - e.g. controlling for SEP, res etc.. sub national level

```{r}
# Urban vs Rural, Ethiopia
ggplot(eth, aes(x = mar, y = mpa))+
  geom_point()+
  facet_wrap(~res)

# Urban vs Rural, India
ggplot(ind, aes(x = mar, y = mpa))+
  geom_point()+
  facet_wrap(~res)

# Urban vs Rural, Nigeria
ggplot(nga, aes(x = mar, y = mpa))+
  geom_point()+
  facet_wrap(~res)

# All the distributions look quite different but in each country, similar

```
